
\begin{center}
\Large
    Generative AI tools can swing from being absolute game-changers to the very bane of my existence. 
    Prof. Jessy Grizzle, 
    Fall 2023
\end{center}

During the writing of this book, from May 2023 until August 2024, I have used both ChatGPT4 (the paid version) and Bard. In general, they were very helpful for
\begin{itemize}
    \item Exploring math ideas in my search for a concise, clear path from Linear Algebra through Calculus, with a heavy emphasis on how Engineers use Calculus in real applications.
    \item What got me started was this interaction with ChatGPT in March of 2023.
    \section*{Integration Problem}

\begin{shaded}
\textbf{User:} 
Find the antiderivative of \(x^2\ln(x)\)
\end{shaded}
\item Hence, ChatCPT is at least as good as I am at integration by parts and likely much better. Now that we can integrate into ChatGPT the Wolfram plugin, it is definitely better than me. Way better! Our students will use these tools as professionals. I'd be remiss to pretend otherwise.
\item ChatGPT is also a decent programmer. Throughout the writing of this book, I've learned to give ChatGPT a code stub, along with basic instructions for what I wanted the final code to do, and it would complete it. With a few tweaks on my part, a bit of back-and-forth between ChatGPT and me, \textbf{four times out of five}, I'd have a quality piece of code that I could use in the book.
\item \textbf{It's the other one time out of five that drives me crazy, with hours lost going in a vicious circle}. More ranting to do. 


\begin{shaded}
\textbf{ChatGPT:} 
Certainly! Let's solve the problem using integration by parts. \\

Integration by parts is given by:
\[
\int u \, dv = uv - \int v \, du
\]

For the integral:
\[
\int x^2 \ln(x) \, dx
\]

Let's choose:
\[
u = \ln(x) \implies du = \frac{1}{x} \, dx
\]
\[
dv = x^2 \, dx \implies v = \frac{x^3}{3}
\]

Now, plugging into the integration by parts formula:
\[
\int x^2 \ln(x) \, dx = x^3 \ln(x) - \int \frac{x^3}{3} \cdot \frac{1}{x} \, dx
\]
\[
= x^3 \ln(x) - \int \frac{x^2}{3} \, dx
\]
\[
= x^3 \ln(x) - \frac{x^3}{9} + C
\]

Where \( C \) is the constant of integration.
\end{shaded}

\textbf{Date:} March 23, 2023 \\
\textbf{Last Training Date of ChatGPT:} January 2022.
\end{itemize}





